\input{../report.input.tex}
\chapter{APT}
$\text{Errata: "[...]dpkg, is one of the most powerful"}^{\text{citation needed}}$
\section{Exercise 1: Review and preparation}
\subsection{1-1 Read the documentation for apt-get and answer the following questions.}
\subsubsection{a) Which apt-get sub-command is used to install new packages.}
\verb=# apt-get install <package>=

\subsubsection{b) Which apt-get sub-command is used to remove packages.}
\verb=# apt-get remove|purge <package>=

\subsubsection{c) What does apt-get update do.}
\verb=update= gets the latest index files from the repositories.

\subsubsection{d) What does apt-get dist-upgrade do.}
It upgrades all explicitly installed packages and (hopefully) their dependencies.

\subsection{1-2 Read the documentation for dpkg and answer the following questions.}
\subsubsection{a) What does dpkg --get-selections do.}
Prints a list of installed packages, unless a search pattern is given, in which case not installed in the indexes also are included (well, the ones that matches the search pattern).

\subsubsection{b) What does dpkg --purge vim do.}
Removes the package and its config files.

\subsubsection{c) What does dpkg -L bind9 do.Exercise 1: Review and preparation}
c) Lists files installed by \verb=bind9=.
% 2-1
% # /etc/apt/sources.list
% deb http://ftp.se.debian.org/debian squeeze
% deb http://security.debian.org/debian squeeze updates
%
% deb-src <url> <dist> <component>

% 3-1,2
% # apt-get install aptitude hping

% 4-1
% # apt-get remove hping

\section{Exercise 6: Review of aptitude}
\subsection{6-1 What command in aptitude will cause the package list to contain only those packages whose names contain a specific word?}
\verb=/ <word>=

\subsubsection{6-2 What command in aptitude lets you search the package list without hiding anything?}
\verb=l <word>=

\subsubsection{6-3 What command marks a package for installation? For deletion?}
\begin{verbatim}# aptitude install <package>
# aptitude remove <package>\end{verbatim}
('+'and '-' keys)

\subsubsection{6-4 What command causes all selected actions (installation, deletion, and upgrades) to take place.}
\verb=# aptitude -i=\\
('g'-key)

\subsubsection{6-5 If there is a B in the leftmost column, what does that mean.}
The package has broken dependencies.

\end{document}
